\qns{Inverse of a Matrix-Matrix Product}

\textbf{Learning Goal:} This problem aims to familiarize you with the properties of inverse and related proof techniques.

\meta{
\begin{itemize}
\item Stress the importance of ordering in matrix multiplication! For instance, students may be inclined to multiply the outer matrices together. 
\item For these types of problems, the trick is to see how you can reduce a complicated matrix expression by identifying the pairs that result in an identity matrix.

\item Maybe mention that, although unnecessary, another valid method of proof would be to construct the inverse, for example:

\begin{itemize}
\item[$\blacksquare$] $\mathbf{A}\mathbf{B}(\mathbf{A}\mathbf{B}^{-1}) = 1 $

\item[$\blacksquare$] $\mathbf{A}^{-1}\mathbf{A}\mathbf{B}(\mathbf{A}\mathbf{B}^{-1}) = \mathbf{A}^{-1}$

\item[$\blacksquare$] $\mathbf{B}(\mathbf{A}\mathbf{B}^{-1}) = \mathbf{A}^{-1}$

\item[$\blacksquare$] $\mathbf{B}^{-1}\mathbf{B}(\mathbf{A}\mathbf{B}^{-1}) =\mathbf{B}^{-1} \mathbf{A}^{-1}$

\item[$\blacksquare$] $(\mathbf{A}\mathbf{B})^{-1} =\mathbf{B}^{-1} \mathbf{A}^{-1}$

\end{itemize}


\end{itemize}
}

Prove that if a matrix-matrix product $\mathbf{A}\mathbf{B}$ is invertible,  the inverse will be equal to $\mathbf{B}^{-1}\mathbf{A}^{-1}$. Please see \notes{Note 6: subsection 6.1.1} for properties of inverse. \\ \\
HINT: We start again with what we know.
Since $\mathbf{A}\mathbf{B}$ is invertible, we know that an inverse exists, i.e.
\begin{align*}
(\mathbf{A}\mathbf{B})(\mathbf{A}\mathbf{B})^{-1}&=\mathbf{I}\\
(\mathbf{A}\mathbf{B})^{-1}(\mathbf{A}\mathbf{B})&=\mathbf{I}
\end{align*}

%Prove that if a matrix $\mathbf{A}$ is not invertible,  $\mathbf{B}\mathbf{A}$ will not be invertible.

\ans{

Now $\mathbf{B}^{-1}\mathbf{A}^{-1}$ will be an inverse of $\mathbf{A}\mathbf{B}$, if and only if the following relation is satisfied: 
\begin{align*}
(\mathbf{B}^{-1}\mathbf{A}^{-1})(\mathbf{A}\mathbf{B})=(\mathbf{A}\mathbf{B})(\mathbf{B}^{-1}\mathbf{A}^{-1})=\mathbf{I}. 
\end{align*}

For this proof, we are going to evaluate both $(\mathbf{B}^{-1}\mathbf{A}^{-1})(\mathbf{A}\mathbf{B})$ and $(\mathbf{A}\mathbf{B})(\mathbf{B}^{-1}\mathbf{A}^{-1})$ and see if they are equal to the identity matrix. First let us start with:
\begin{align*}
(\mathbf{B}^{-1}\mathbf{A}^{-1})(\mathbf{A}\mathbf{B})
&=\mathbf{B}^{-1}\mathbf{A}^{-1}\mathbf{A}\mathbf{B}\\
&=\mathbf{B}^{-1}(\mathbf{A}^{-1}\mathbf{A})\mathbf{B}\\
&=\mathbf{B}^{-1}\mathbf{I}\mathbf{B}\\
&=\mathbf{B}^{-1}(\mathbf{I}\mathbf{B})\\
&=\mathbf{B}^{-1}\mathbf{B}\\
&=\mathbf{I}
\end{align*}
Then we evaluate:
\begin{align*}
(\mathbf{A}\mathbf{B})(\mathbf{B}^{-1}\mathbf{A}^{-1})
&=\mathbf{A}\mathbf{B}\mathbf{B}^{-1}\mathbf{A}^{-1}\\
&=\mathbf{A}(\mathbf{B}\mathbf{B}^{-1})\mathbf{A}^{-1}\\
&=\mathbf{A}(\mathbf{I})\mathbf{A}^{-1}\\
&=\mathbf{A}\mathbf{A}^{-1}\\
&=\mathbf{I}
\end{align*}
Since both of expressions are equal to identity matrix and the inverse for a matrix is unique, we can decide that
 $(\mathbf{A}\mathbf{B})^{-1}=\mathbf{B}^{-1}\mathbf{A}^{-1}$ is the inverse of $\mathbf{A}\mathbf{B}$. So the proof is complete.

}