\qns{Bode Plots for Filters}
\qcontributor{Ramsey Mardini}
\qcontributor{Naomi Sagan}
\qcontributor{Ryan Koh}

\begin{enumerate}

\qitem First, consider the following transfer function of a Low-Pass Filter: $H(\omega) = \frac{1}{j\omega CR + 1}$
    \begin{enumerate}
        \qitem What is its cutoff frequency?

        \sol{TBD}
        \qitem Sketch its phase and magnitude.

        \sol{TBD}
    \end{enumerate}

\qitem Now consider the transfer function of a High-Pass Filter: $H(\omega) = \frac{j\omega CR}{j\omega CR + 1}$
    \begin{enumerate}
        \qitem What is its cutoff frequency?

        \sol{TBD}
        \qitem Sketch its phase and magnitude.

        \sol{TBD}
    \end{enumerate}

\qitem What happens if we cascade these two filters together with a unity-gain buffer in between them? Consider the resulting transfer function: 
        $H(\omega) = \frac{1}{j\omega CR + 1} \cdot \frac{j\omega CR}{j\omega CR + 1}$
    \begin{enumerate}
        \qitem What are its cutoff frequencies?

        \sol{TBD}
        \qitem Sketch its phase and magnitude. \textit{Hint: How can we combine the plots of the individual filters together?}

        \sol{TBD}
    \end{enumerate}
Now lets take a look at two more concepts, \textbf{bandwidth} and \textbf{Q-factor}. The bandwidth of a resonance
circuit is the difference between its upper and lower cutoff frequencies. It can also be expressed as \textit{resonance frequency} divided by its \textit{Q-factor}:

\begin{align}
    bandwidth = \Delta \omega = f_{high} - f_{low} = \frac{\omega_n}{Q}
\end{align}

The Q-factor, or quality factor, of a circuit is the ratio of power stored to power dissipated, but can also be more simply thought of as the measure of how good a circuit is,
where a higher Q-factor is often more desirable. This can be expressed mathematically as:

\begin{align}
    Q = \frac{\omega_{n}}{\Delta \omega}
\end{align}

    \begin{enumerate}[resume]
        \qitem Find its bandwidth.

        \sol{TBD}
        \qitem What is its Q-factor? I.e.: $\frac{\omega_{n}}{\Delta \omega}$, where $\omega_{n}$ is the center frequency of the band.
        
        \sol{TBD}
    \end{enumerate}
\end{enumerate}
