\qns{Weighted minimum norm} {\bfseries (Spring 2019 Final)}
\qcontributor{Nikhil Shinde}

You saw in lecture in the context of open-loop control, how we
consider problems in which we have a wide
matrix $A$ and solve $A \vec{x} = \vec{y}$ such that $\vec{x}$
is a minimum norm solution:  
\begin{align*}
\norm{\vec{x}} \leq \norm{\vec{z}} 
\end{align*}
for all $\vec{z}$ such that $A \vec{z} = \vec{y}$. You have worked out how to compute the appropriate ``pseudo-inverse'' for
such wide matrices.

But what if you weren't interested in just the norm of $\vec{x}$? What
if you instead cared about minimizing the norm of a linear transformation $C\vec{x}$? For example, suppose that controls were more or less costly
at different times. 

The problem can be written out mathematically as:

Given a wide matrix  $A$ and a matrix $C$ find $\vec{x}$ such that $A \vec{x} = \vec{y}$ and 
$\norm{C\vec{x}} \leq \norm{C\vec{z}} $ for all 
$\vec{z}$ such that $A \vec{z} = \vec{y}$.


{\bf(10 pts)} Let's start with the case of $C$ being invertible. {\bfseries Solve
  this problem (i.e. find the optimal $\vec{x}$ with the minimum $\|C \vec{x}\|$) for the specific
  matrices and $\vec{y}$ given below. Show your work.}

It is fine to leave your answer as an explicit product of
matrices and vectors.

{\em (HINT: You might want to change variables to solve this
  problem. Don't forget to change back!)} 

\begin{align*}
A = \begin{bmatrix}
1 & 0 & 0 \\
0& 1 & 1
\end{bmatrix}, \quad
C = \begin{bmatrix}
0 & 0 & 2\\
0 & 1 & 0 \\
2 & 0 & 0
\end{bmatrix}, \quad 
\vec{y} = \begin{bmatrix}
2 \\
1
\end{bmatrix}
\end{align*}

For convenience, $C^{-1} = \begin{bmatrix}
0 & 0 & 0.5\\
0 & 1 & 0 \\
0.5 & 0 & 0
\end{bmatrix}$ and you are also given some SVDs on the following page.

\begin{align}
A &= (U_A = \begin{bmatrix}
0& 1  \\
1 & 0
\end{bmatrix})(
\Sigma_A = \begin{bmatrix}
\sqrt{2}  &  0  & 0\\
0 & 1 & 0
\end{bmatrix})(
V^T_A = \begin{bmatrix}
0& \frac{1}{\sqrt{2}}& \frac{1}{\sqrt{2}}  \\
1 & 0 & 0 \\
0& -\frac{1}{\sqrt{2}}& \frac{1}{\sqrt{2}}  
\end{bmatrix}) \\
C &= 
(U_C = \begin{bmatrix}
-1 & 0 & 0  \\
0 & 0 & 1 \\
0 & -1 & 0
\end{bmatrix})
(\Sigma_C = \begin{bmatrix}
2& 0 & 0  \\
0 & 2 & 0 \\
0 & 0 & 1
\end{bmatrix})
(V^T_C = \begin{bmatrix}
0 & 0 & -1  \\
-1 & 0 & 0 \\
0 & 1 & 0
\end{bmatrix}) \\
AC &= 
(U_{AC} = \begin{bmatrix}
0 & 1 \\
1 & 0
\end{bmatrix}
(\Sigma_{AC} = \begin{bmatrix}
\sqrt{5}  &   0 & 0\\
0 & 2 & 0  \\
\end{bmatrix})
(V^T_{AC} = \begin{bmatrix}
\frac{2}{\sqrt{5}}&\frac{1}{\sqrt{5}} & 0 \\
0& 0 & 1 \\
-\frac{1}{\sqrt{5}} & \frac{2}{\sqrt{5}} & 0
\end{bmatrix}) \\
AC^{-1} &= 
(U_{AC^{-1}} = \begin{bmatrix}
0 & 1 \\
1 & 0
\end{bmatrix})
(\Sigma_{AC^{-1}} = \begin{bmatrix}
\frac{\sqrt{5}}{2}  &   0 & 0\\
0 & 0.5 & 0  \\
\end{bmatrix})
(V^T_{AC^{-1}} = \begin{bmatrix}
\frac{1}{\sqrt{5}}&\frac{2}{\sqrt{5}} & 0 \\
0 & 0 & 1 \\
-\frac{2}{\sqrt{5}} & \frac{1}{\sqrt{5}} & 0
\end{bmatrix})
\end{align}

\sol{ 
We already know to find the minimum-norm solution for the system of equations
$$A\vec{x} = \vec{y}$$
using the \textbf{Moore-Penrose Pseudoinverse}. \\

To use that knowledge, we want to first change variables to the C-basis:
$$\widetilde{x} = C\vec{x}$$
Given an arbitrary vector 
$$\vec{p} = C\vec{z}$$
we get the new constraint $\norm{\widetilde{x}}
\leq \norm{\vec{p}}$ for any vector $\vec{p}$ that satisfies
something. What is this something? \\
\newline

Originally, we had $A \vec{x} = \vec{y}$ and, in the new basis, we have $\vec{x} = C^{-1} \widetilde{x}$. 
So, the constraint that needs to be satisfied is 
$$AC^{-1} \widetilde{x} =
\vec{y}$$ 

This is a min-norm problem where the matrix multiplying the vector is $AC^{-1}$. 
To solve this, we find the Moore-Penrose psuedoinverse of $AC^{-1}$:
\begin{align}
\widetilde{x} =
  V_{compact, AC^{-1}} \Sigma_{compact, AC^{-1}}^{-1} U_{AC^{-1}}^T\vec{y}
\end{align}

In this case, we use the compact form of the SVD so the $\Sigma$ matrix is square and therefore invertible.
The compact SVD of $AC^{-1}$ is as follows:
\begin{align}
AC^{-1} &= 
(U = \begin{bmatrix}
0 & 1 \\
1 & 0
\end{bmatrix})
(\Sigma = \begin{bmatrix}
\frac{\sqrt{5}}{2}  &   0 \\
0 & 0.5  
\end{bmatrix})
(V^T = \begin{bmatrix}
\frac{1}{\sqrt{5}}&\frac{2}{\sqrt{5}} & 0 \\
0 & 0 & 1 
\end{bmatrix}).
\end{align}

Calculating this out, we get that $\widetilde{x}
= \begin{bmatrix}\frac{2}{5} \\ \frac{4}{5} \\ 4\end{bmatrix}$. 

However since the original question was to find $\vec{x}$ we have one more substitution to arrive at our final answer:
\begin{align}
\vec{x} &= C^{-1}\widetilde{x} = C^{-1}   V_\Sigma^{-1}U^T\vec{y}
\\
&= \begin{bmatrix}
2 \\
0.8 \\
0.2
\end{bmatrix}.
\end{align}
}