% Author: Nareauphol (Gavin) Liu

\qns{Upper Triangularization}

\qcontributor{Nareauphol Liu}

Recall that before we solved the system of differential equation $\frac{d}{dt}\vec{x}(t) = A\vec{x}(t) + \vec{b}u(t)$ using 
the change of basis $V$ which contained the eigenvectors of A. The result of the transformation showed that $V^{-1}AV = D$ 
became the diagonal matrix 
\[
  D = \begin{bmatrix}
    \lambda_1 & ... & 0 \\
    \vdots & \ddots & \vdots \\
    0 & ... & \lambda_n
  \end{bmatrix}
\]
where $\lambda_1, ..., \lambda_n$ are the eigenvalues of $A$

However, the underlying assumption for this transformation is that our matrix $A$ was diagonalizable. In such a case that 
our matrix $A$ is not diagonalizable, we can still solve the system of differential equation using upper triangularization by 
solving it "bottom-up" using backwards substitution. 

\begin{enumerate}
  \qitem Give an example of a $2\times2$ matrix that is not diagonalizable. Explain why we can't use our original trick $D = V^{-1}AV$.

  \ws{
  \vspace{150px}
  }

  \sol {
    Any answer can suffice but the most common example of a non-diagonalizable matrix is 
    \[
    A = \begin{bmatrix}
        1 & 1 \\
        0 & 1
    \end{bmatrix}
    \]

    where there is only one repeated eigenvalue $\lambda = 1$ and the corresponding eigenvector 
    $\begin{bmatrix}
        1 \\
        0 
    \end{bmatrix}$
    We can no longer use $D = V^{-1}AV$ becomes we need 2 eigenvectors to construct the $V$ basis but we only have 1 eigenvector.
  }
   

  \qitem For this given upper triangular system of differential equations
  \[
    \frac{d}{dt}\vec{x}(t) = \begin{bmatrix}
        \lambda_1 & a & ... & A_{1,n-1} & A_{1, n} \\
        0 & \lambda_2 & ...  & A_{2,n-1} & A_{2,n}\\
        \vdots & 0 & \ddots & \lambda_{n-1} & A_{n-1,n} \\
        0 & 0 & ... & 0 & \lambda_n
        \end{bmatrix}
        \vec{x}(t)
    \]
  Explain how we can use upper triangularization to solve the system of differential equation.\\
  \textit{Hint: Which differential equation can we solve immediately? Can we use that solution to backward substitute in anyway?}

    \sol {
   
    Notice that we can solve the differential of the the bottom most row 
    \[ 
    \frac{d}{dt}x_n(t) = \lambda_nx(t) 
    \] 

    to be $x_n(t) = x_n(0)e^{\lambda_n t}$. Notice that in the row above it, the differential equation is at the form 
    \[ 
    \frac{d}{dt}x_{n-1}(t) = \lambda_{n-1}x(t) + A_{n-1,n}x_n(t). 
    \]
    
    Since we've solved for $x_n(t)$ previously, we can simply plug 
    in the solution at the $n-1$ row and our differential equation becomes of the form $\frac{d}{dt}x(t) = \lambda x(t) + u(t)$ 
    which we know how to solve using the integral solution. We can repeat this process and solve each differential equation 
    at the each row bottom-up. This is what "backwards substitution" means. 
  }

  \ws{
  \vspace{150px}
  }


  \qitem We will now perform upper triangularization on a non-diagonalizable $2\times2$ matrix. 
  \[
  A = \begin{bmatrix}
      1 & 1 \\
      -1 & 3 \\
  \end{bmatrix}
  \]Show that 
    $\vec{u_1} = \begin{bmatrix}
        1 \\
        1 
    \end{bmatrix}$ is an eigenvector of $A$ and find its corresponding eigenvalue. 
    \ws{
    \vspace{150px}
    }

  \sol {
    Recall that $A\vec{v} = \lambda \vec{v}$ for any eigenvector-eigenvalue pair. We can simply multiply $A$ and $\vec{u_1}$ to get 
    $A\vec{u_1} = \begin{bmatrix}
        2 \\
        2 
    \end{bmatrix}$
    and clearly we can see that corresponding eigenvalue is $\lambda = 2$
  }
  \qitem Using this eigenvector and some appropriate basis vectors of $R^2$, find an orthonormal basis $U$ for the column space of $A$ 
  using the Gram-Schmidt algorithm. What do you notice about running the Gram-Schmidt algorithm with more vectors that needs
  to span a certain subspace? How does Gram-Schmidt handle this? 

  
  \ws{
  \vspace{150px}
  }
  \meta {
      If the students are still confused about Gram-Schmidt, this is the perfect place to go through the algorithm again in 
      great detail since it is essential to upper triangularization. Also, please show Anish's video on the Gram-Schmidt algorithm 
      pictorically \underline{\href{https://www.youtube.com/watch?v=79Ss_HkwthE}{here}}
  }

  \sol {
      
    We can find an orthonormal basis for any matrix using the Gram-Schmidt algorithm by filling in the basis vectors of $R^3$
    Thus we simply run the Gram-Schmidt algorithm on the columns of the matrix below 
    \[
    U = \begin{bmatrix}
        1 & 1 & 0 \\
        1 & 0 & 1
    \end{bmatrix}
    \]

    Running the Gram-Schmidt algorithm gets us 

    $\vec{u_1} = \begin{bmatrix}
        \sqrt{2}/2 \\
        \sqrt{2}/2  
    \end{bmatrix}$ since the norm of this vector is $\sqrt{2}$


    $\vec{u_2} = \vec{s_2} - proj_{\vec{u_1}}\vec{s_2} = \begin{bmatrix}
        1/2 \\
        -1/2
    \end{bmatrix}$ 

    Now we normalize $\vec{u_2}$ to be 
    $
    \begin{bmatrix}
        \sqrt{2}/2\\
        -\sqrt{2}/2
    \end{bmatrix}
    $
    
    $\vec{u_3} = \vec{s_3} - proj_{\vec{u_1}}\vec{s_3} - proj_{\vec{u_2}}\vec{s_3} = 
    \begin{bmatrix}
        0 \\
        0 
    \end{bmatrix}$ \\

    Since we added more vectors than we need to span $R^2$, Gram-Schmidt automatically will output one zero vector and we simply 
    discard this vector. This is because we are inputting one more vector into our basis than we need. The zero vector implies that
    the extra vector is linearly dependent to the other vectors. Intuitively this makes sense since if we put $n+1$ vectors into 
    Gram-Schmidt to try to span $R^N$, we should only need $n$ vectors, and thus it will always output one zero vector. 
    Thus our orthonormal basis $U$ for $A$ is 
    \[
    U = \begin{bmatrix}
    \sqrt{2}/2 &  \sqrt{2}/2\\
    \sqrt{2}/2 &  -\sqrt{2}/2
    \end{bmatrix}
    \]

    }

    \qitem The next step is to compute $Q = U^TAU$. Compute $Q$. What do you notice about the matrix 
    $Q$? What is interesting about the diagonals of $Q$?
    \ws{
    \vspace{150px}
    }

    \sol{
        \[
        Q =  \begin{bmatrix}
            \sqrt{2}/2 &  \sqrt{2}/2\\
            \sqrt{2}/2 &  -\sqrt{2}/2
            \end{bmatrix}^T 
            \begin{bmatrix}
                1 & 1 \\
                -1 & 3 \\
            \end{bmatrix}
            \begin{bmatrix}
                \sqrt{2}/2 &  \sqrt{2}/2\\
                \sqrt{2}/2 &  -\sqrt{2}/2
            \end{bmatrix} =
        \begin{bmatrix}
            2 &  2\\
            0 &  2
        \end{bmatrix}
        \]

        Notice that the matrix is upper triangular and the diagonals of $Q$ corresponds to the eigenvalues of $A$
    }


    \qitem Notice that we have computed $U^TAU = Q = \begin{bmatrix}
        \lambda_1 &  2\\
        0 &  \lambda_2
    \end{bmatrix}$. Show that you can write $A = U \begin{bmatrix}
        \lambda_1 &  2\\
        0 &  \lambda_2
    \end{bmatrix} U^T$.
    \textit{Hint: What is special about the matrix $U$?}

    \ws{
    \vspace{150px}
    }

    \meta {
        It might be helpful to first try to get them to show that $U^T = U^{-1}$ for any orthonormal matrix $U$. From there, 
        have them convince themselves that $UU^T = U^TU = I$
    }

    \sol {
        From earlier, we can write $Q =  U^T A U$. Since $U$ is an orthonormal matrix, we can show that 
        \[
        Q =  U^T A U 
        \]

        Multiply both sides with $U$ to get
        \[
            UQ = UU^TAU = AU 
        \]

        Multiply both sides with $U^T$ to get
        \[ 
            UQU^T = AUU^T = A
        \]

        Thus we finally have
        \[
        A =  UQU^T = U \begin{bmatrix}
            \lambda_1 &  2\\
            0 &  \lambda_2 \end{bmatrix}  U^T
        \]
    }

    \qitem Now let's extend this to a $3\times3$ matrix. Play around with the Jupyter Notebook to see the process of 
    upper triangularizing a $3\times3$ matrix.  

    \ws{
    \vspace{150px}
    }

    \meta {
        It is advised and probably more helpful if you go through the discussion worksheet and solutions here 
        \underline{\href{https://www.eecs16b.org/discussion/dis09A.pdf}{here}} and walk them through 
        the Jupyter Notebook Datahub link carefully and show that you need to run
        Gram-Schmidt twice in order to construct the $U$ basis which upper triangularizes the $3\times3$ matrix $A$ 
    }

    \sol {
        On Jupyter Notebook 
    }

    % \qitem What should we pick our basis $U$ such that we can upper triangularize our matrix $A$?

    % \ws{
    % \vspace{150px}
    % }

    % \sol {
    %     We should pick $U = \begin{bmatrix}
    %         \sqrt{2}/2 &  \sqrt{2}/2\\
    %         \sqrt{2}/2 &  -\sqrt{2}/2
    %         \end{bmatrix}$ in order to upper triangularize our matrix as we have shown in part (e)
    % }

    \qitem Notice that we got lucky that we only needed to run Gram-Schmidt once to find a basis which 
    upper triangularizes our $2\times2$ matrix $A$. For higher dimension matrices like the $3\times3$ matrix in the Jupyter Notebook, 
    however, this is not the case  and multiple iterations of part (d)-(f) is required to construct such basis. 
    For a general matrix $n\times n$,  how many times do we need to perform Gram-Schmidt in order to 
    construct a basis that upper triangularizes $A$? 

    \textit{Hint: In the $3\times3$ matrix example in the Jupyter Notebook, how many times was Gram-Schmidt algorithm ran?
    Why did it run that many times?}

    \ws{
    \vspace{150px}
    }

    \sol {
        We need to run it $n-1$ times to construct an orthonormal basis to upper triangularize our matrix $A$. This is because 
        for higher dimensional matrix, running Gram-Schmidt once doesn't guarantee
        an upper triangular matrix when multiplying $U^TAU$ because you'll produce a smaller submatrix which isn't upper
        triangular. Thus we will need to keep repeating the process on smaller submatrices  until the $2\times2$ case since the smaller submatrix of that is a 1x1 matrix, which is a scalar. 
    }

    \qitem In your own words, write down the general algorithm for upper triangularizing an $n\times n$ matrix $A$. 

    \meta {
        Feel free to add your own touch to the algorithm. The solution is the general guideline to what the algorithm does. 
        Also remind them that in an exam setting, students won't get any matrix bigger than a $3\times3$ matrix and homework
        will have a numerical problem for them to do which would solidify their understanding of upper triangularization
    }

    \ws{
    \vspace{150px}
    }

    \sol {
        Step 1: Find an eigenvalue-eigenvector pair for $A$. \\
        Step 2: Using the appropriate basis vectors and the eigenvector from Step 1, 
        run Gram-Schmidt  to construct an orthonormal basis. Don't forget to throw out 
        $\vec{0}$ since we have one extra vector everytime we do this step. \\
        Step 3: Repeat Step 1 and 2 on the smaller submatrix until the $2\times2$ case. \\
        Step 4: Using what Gram-Schmidt outputs, construct the orthonormal basis $U$.
    }

    

  
\end{enumerate}
