% Author: Naomi Sagan, Ryan Koh
% Emails: naomi.sagan@berkeley.edu, ryan_koh@berkeley.edu

\qns{Intro to Discrete-Time Systems}

\usetikzlibrary{calc, automata, chains, arrows.meta}

Students are studying for the EECS16C exam, and the flow of students from Moffit to Taejin's Office Hours (OH) can be represented as the following: \newline

\begin{align*}
    \begin{tikzpicture}
        \node[circle, draw=black, minimum size=2cm] (Moffit) {Moffit};
        \node[circle, draw=black, minimum size=2cm] (OH) [right=of Moffit] {OH};
        \draw [->] (Moffit.north east) to [bend left=15]  node[above] {0.2}  (OH.north west);
        \path (Moffit) edge [loop above] node {0.8} (Moffit);
        \path (OH) edge [loop above] node {0.4} (OH);
        \node (u_moffit) [left=of Moffit] {$0.5u[t]$};
        \draw [->] (u_moffit.east) to []  node[left] {}  (Moffit.west);
        \node (u_OH) [right=of OH] {$0.5u[t]$};
        \draw [->] (u_OH.west) to []  node[right] {}  (OH.east);
        \draw [->] (OH.south west) to [bend left=15] node[below] {0.6} (Moffit.south east);
    \end{tikzpicture}
\end{align*}

where $u[t]$ is the number of students that start to study at timestep $t$. (i.e.: $u[t]$ is the input to the system)

\begin{enumerate}
    \qitem Let our state variables be represented by $x_1$ and $x_2$. \textbf{Explain in your own words what $x_1$ and $x_2$ are.}
        \sol{
            We can represent our state by the number of students in Moffit ($x_1$) and the number of students in Taejin's office hours ($x_2$).
        }
    \vspace{2.5cm}
    \qitem Represent the flow of students between the two states as a matrix-vector discrete time system:
    \begin{align*}
        \vec{x}[t+1] =  A\vec{x}[t] + \vec{b}u[t]\\
    \end{align*}
    \textbf{Find matrix $A$ and vector $\vec{b}$.} \\
    \sol{
        \begin{align*}
            \begin{bmatrix}
                x_1[t + 1] \\ x_2[t + 1]
            \end{bmatrix} =
            \begin{bmatrix}
                0.8 & 0.6 \\
                0.2 & 0.4
            \end{bmatrix}
            \begin{bmatrix}
                x_1[t] \\ x_2[t]
            \end{bmatrix} +
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} u[t]
        \end{align*}
    }
    \vspace{8cm}
    \qitem Let $\vec{x}[0] = \vec{0}$, and $u[t] = 10$ for all values of $t$. \textbf{What is $\vec{x}[1]$? What is $\vec{x}[2]$?} \textbf{What is $||\vec{x}||$ as $t\rightarrow\infty$? Does that make sense in the context of our problem?}

    \sol{
        \begin{align*}
            \vec{x}[1] = A \vec{0} + 10
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} =
            \begin{bmatrix}
                5 \\ 5
            \end{bmatrix} \\
            \vec{x}[2] = A
            \begin{bmatrix}
                5 \\ 5
            \end{bmatrix} + 10
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} =
            \begin{bmatrix}
                0.8 & 0.6 \\
                0.2 & 0.4
            \end{bmatrix}
            \begin{bmatrix}
                5 \\ 5
            \end{bmatrix} +
            \begin{bmatrix}
                5 \\ 5
            \end{bmatrix} =
            \begin{bmatrix}
                12 \\ 8
            \end{bmatrix}
        \end{align*}

        The input to the system is always positive, and no students ever leave the system, so the magnitude of the state keeps on growing. So,
        \begin{align*}
            \lim_{t \to \infty} ||\vec{x}[t]|| = \infty
        \end{align*}
        This does not make sense in the context of our problem because it is impossible to have an infinite amount of students.
    }

    \vspace{8cm}
    \qitem Let $\vec{x}[0] = \begin{bmatrix} 50 \\ 50 \end{bmatrix}$, and $u[t] = -4$ for all values of $t$. \textbf{What is $\vec{x}[1]$? What is $\vec{x}[2]$?} \textbf{What is $||\vec{x}||$ as $t\rightarrow\infty$? What sign are the elements of $\vec{x}$? Does that make sense in the context of our problem?}

\sol{
        \begin{align*}
            \vec{x}[1] = A
            \begin{bmatrix}
                50 \\ 50
            \end{bmatrix} - 4
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} =
            \begin{bmatrix}
                0.8 & 0.6 \\
                0.2 & 0.4
            \end{bmatrix}
            \begin{bmatrix}
                50 \\ 50
            \end{bmatrix} -
            \begin{bmatrix}
                2 \\ 2
            \end{bmatrix} =
            \begin{bmatrix}
                68 \\ 28
            \end{bmatrix} \\
            \vec{x}[2] = A
            \begin{bmatrix}
                68 \\ 28
            \end{bmatrix} - 4
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} =
            \begin{bmatrix}
                0.8 & 0.6 \\
                0.2 & 0.4
            \end{bmatrix}
            \begin{bmatrix}
                68 \\ 28
            \end{bmatrix} -
            \begin{bmatrix}
                2 \\ 2
            \end{bmatrix}  =
            \begin{bmatrix} 69.2 \\ 22.8 \end{bmatrix}
        \end{align*}
        \textit{Note: it's ok that the state variables are not whole numbers! Think of them as the average number of students in Moffit and Taejin's OH.}
        \newline
        The input to the system is always negative, and no students are ever added to the system, so the magnitude of the state keeps on decreasing linearly. So, \textbf{the sign of the elements of $\vec{x}$ is negative} and
        \begin{align*}
            \lim_{t \to \infty} ||\vec{x}[t]|| = \infty
        \end{align*}
        This does not make sense in the context of our problem because there is an infinitely negative amount of students in our system.
    }

    \vspace{12cm}
    \qitem Let $\vec{x}[0] = \vec{0}$, $u[0] = 16$, and $u[t>0] = 0$. \textbf{What is $\vec{x}[1]$? What is $\vec{x}[2]$?} \textbf{What is the largest $||\vec{x}||$ can get as $t\rightarrow\infty$? Does that make sense in the context of our problem?}

    \sol {
        \begin{align*}
            \vec{x}[1] = A \vec{0} + 16
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} =
            \begin{bmatrix}
                8 \\ 8
            \end{bmatrix} \\
            \vec{x}[2] = A
            \begin{bmatrix}
                8 \\ 8
            \end{bmatrix} + 0
            \begin{bmatrix}
                0.5 \\ 0.5
            \end{bmatrix} =
            \begin{bmatrix}
                0.8 & 0.6 \\
                0.2 & 0.4
            \end{bmatrix}
            \begin{bmatrix}
                8 \\ 8
            \end{bmatrix} =
            \begin{bmatrix} 11.2 \\ 4.8 \end{bmatrix}
        \end{align*}

        The input to the system is 0 after the first time step, and the sum of students in the two states stays the same each timestep, so, by the triangle inequality, the magnitude of $\vec{x}$ will remain less than or equal to the sum of the students in the two states after the first timestep.
        \begin{align*}
            \lim_{t \to \infty} ||\vec{x}[t]|| \leq 16
        \end{align*}
        This \textit{does} make sense in the context of our problem because it is a finite, positive number.
    }

\end{enumerate}
